\bf Kaggle-in-class Data Challenges Can Boost Student Learning

### 2021-01-04 Submission3/Submitted/paper-kaggle-2ndrevision.Rmd
Submission 3 created

# to produce blinded version set to 1
blinded: 1

@@

<!-- Blinded verssion -->
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of AB (Computational statistics and data mining ,MAST90083, denoted as CSDM), and one at CD University (Statistical thinking, ETC2420/5242, denoted as ST).

<!-- Unblinded verssion 
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of Melbourne (MAST90083), and one at Monash University (ETC2420/5242). -->

@@

Statistical Thinking (ST), covers regression, but not classification, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (ST-PG) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the CSDM competition. The 145 undergraduate (ST-UG) students were used as comparison for examining performance of the postgraduate students. Although, it may be surprising at first glance, the undergraduate students provide a reasonable comparison for the graduate students. The class is taught to both cohorts simultaneously. The entry requirements to the Bachelors of Commerce at Monash is high, and these students have strong mathematics backgrounds. In the years prior to this experiment, the undergraduate scores on the final exam are indistinguishable from those of the graduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete.

@@

Creating a new competition is surprisingly easy. Information on setting up a Kaggle InClass challenge is available on the service's web site (https://www.kaggle.com/about/inclass/overview). There is a setup wizard for step-by-step guidance on getting your competition underway. We have created a short video illustrating the steps to establish a new competition, available on the web (\url{https://www.youtube.com/XXXX}).

<!-- Unblinded verssion -->
<!-- https://www.youtube.com/watch?v=tqbps4vq2Mc&t=32s -->

@@

This document was produced in R [@R] with the package knitr [@knitr]. Data cleaning was conducted using tidyr [@tidyr], dplyr [@dplyr] and plots were made with ggplot2 [@ggplot2]. The materials to reproduce the work are available at https://github.com/XXX.

<!--
https://github.com/dicook/paper-quoll. 
-->

>>>>

# to produce blinded version set to 1
blinded: 0

@@

<!-- Blinded verssion 
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of AB (Computational statistics and data mining ,MAST90083, denoted as CSDM), and one at CD University (Statistical thinking, ETC2420/5242, denoted as ST).

<!-- Unblinded verssion -->
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of Melbourne (Computational statistics and data mining ,MAST90083, denoted as CSDM), and one at Monash University (Statistical thinking, ETC2420/5242, denoted as ST).

@@

Statistical Thinking (ST), covers regression, but not classification, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (ST-PG) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the CSDM competition. The 145 undergraduate (ST-UG) students were used as comparison for examining performance of the postgraduate students. Although, it may be surprising at first glance, the undergraduate students provide a reasonable comparison for the graduate students. The class is taught to both cohorts simultaneously. The entry requirements to the Bachelors of Commerce at 
<!-- Blinded verssion 
CD University -->
<!-- Unblinded verssion -->
Monash 
is high, and these students have strong mathematics backgrounds. In the years prior to this experiment, the undergraduate scores on the final exam are indistinguishable from those of the graduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete.

@@

Creating a new competition is surprisingly easy. Information on setting up a Kaggle InClass challenge is available on the service's web site (https://www.kaggle.com/about/inclass/overview). There is a setup wizard for step-by-step guidance on getting your competition underway. We have created a short video illustrating the steps to establish a new competition, available on the web (\url{https://www.youtube.com/watch?v=tqbps4vq2Mc&t=32s}).

<!-- blinded verssion -->
<!-- https://www.youtube.com/XXXX -->

@@

<!-- blinded verssion 
This document was produced in R [@R] with the package knitr [@knitr]. Data cleaning was conducted using tidyr [@tidyr], dplyr [@dplyr] and plots were made with ggplot2 [@ggplot2]. The materials to reproduce the work are available at https://github.com/XXX.
-->

<!-- Unblinded verssion -->
This document was produced in R [@R] with the package knitr [@knitr]. Data cleaning was conducted using tidyr [@tidyr], dplyr [@dplyr] and plots were made with ggplot2 [@ggplot2]. The materials to reproduce the work are available at https://github.com/dicook/paper-quoll.

### 2021-01-04 Submission3/Submitted/paper-kaggle-2ndrevision.tex
Submission 3 created

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

@@

@@

Kaggle \citep{kaggle} is a platform for predictive modelling and
analytics competitions where participants compete to produce the best
predictive model for a given data set. It is well-known for its
competitions (e.g. \citet{darkmatter}), some of which come with rich
monetary prizes (e.g. \citet{heritage}). There are also learning
competitions \citep{kagglelearn}, designed to help novices hone their
data mining skills. Winners are typically expected to share their code,
and occasionally new algorithms have emerged, for example, deep neural
networks \citep{HintonDahi12} and XGBoost \citep{ChenXGBoost}.

@@

A competition, like any other assessment method has its advantages and
disadvantages. It brings the `game' feeling, increases the interest
level among students and motivates for higher performance
(\citet{Shindler09}, p 105). But, it may have negative influence if
constructed poorly. Among the negative influences are increased stress
and anxiety, induced by fearing a low ranking, failure or technology
barriers. In addition, students may invest a disproportionate amount of
time and effort into competition. Despite some received criticism, a
properly set competition can benefit the students greatly. The
competition should be relatively short in duration to avoid consuming
undue energy. Students should be clear about the rules and the goal.
They should be properly rewarded and most important, feel that they have
reasonable chance to win or achieve high mark \citep{Shindler09}.

\citet{Shelley08} raised the need for more quantitative and statistical
analysis of evidence in science education. This paper contributes to
this call by offering statistical analysis of theffects on learning of
classroom data competitions.

@@

The experiment was conducted during Semester 2 2017. Data was collected
during two classes, one at the University of AB (MAST90083), and one at
CD University (ETC2420/5242).

@@

Computational Statistics and Data Mining (MAST90083) is designed for
postgraduate level students with math, statistics, information
technology or actuarial backgrounds. It covers modelling both continuous
(regression) and categorical (classification) response variables. The 63
students were randomized into one of two kaggle competitions, one
focused on regression (R) and the other classification (C). This setup
mimics randomized control trials, which is the 'gold standard, in
experiment design (\citet{Shelley09CH1}, ch.~1). Students built
prediction models and made submissions individually for 16 days, and
then were allowed to form groups to compete for another 7 days.

The reason for this strategy was first to motivate each of the students
to think about modelling and be actively engaged in the competitions
through individual submission. The lecturer allowed participants to
create groups towards the end of the competition to illustrate the
advantages of group work and mixing models. Another motivation for this
stratagem was the university policy, requiring a strategy to assign
students individually in group assignments.

One of the 63 students elected not to take part in the competition, and
another student didn't sit the exam, producing a final sample size of
61.

Statistical Thinking (ETC2420/5242), covers regression, but not
classification, and has a mix of undergraduate and postgraduate
students. Only the 34 postgraduate (5242) students were required to
participate in the kaggle competition, and competed in the regression
(R) challenge. This was run independently from the MAST90083
competition. The 145 undergraduate (2420) students were used as controls
for examining performance of the postgraduate students. Although, it may
be surprising at first glance, the undergraduate students provide a
reasonable control for the graduate students. The class is taught to
both cohorts simultaneously. The entry requirements to the Bachelors of
Commerce at Monash is high, and these students have strong mathematics
backgrounds. In the years prior to this experiment, the undergraduate
scores on the final exam are indistinguishable from those of the
graduate students. The competition ran for one month. Students formed
their own teams of 2-4 members to compete.

@@

Figure \ref{fig:MAST90083} shows the data collected in MAST90083.
Performance is plotted against type of question, separately for the
competition they completed. The difference in median scores indicates
performance improvement. Students who completed the classification
competition (left) performed relatively better on the classification
questions than the regression questions in the final exam. Conversely,
students who participated in the regression competition performed
relatively better on the regression questions.

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/Melb-1.pdf}
\caption{\label{fig:MAST90083} Boxplots of performance on regression and
classification questions in the final exam, by type of data competition
completed in MAST90083. Students generally performed better on the
questions corresponding to the competition they participated in.}
\end{figure}

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|r|r|}\hline
Question Set & Median difference & Permutation $p$-value \\\hline
Classification & 0.250 & 0.033\\
Regression & 0.104 & 0.000\\\hline
\end{tabular}
\caption{Comparison of median difference in performance by competition group, for MAST90083 students, using permutation tests. Both sets of medians are significantly different, indicating improved scores for questions on the topic related to the Kaggle competition.}
\end{center}
\label{tab:Melb_Perm}
\end{table}

Table 1 shows the results of permutation testing of median difference
between the groups. Generally the results support the competition
improved performance. Students who participated in the kaggle challenge
for classification scored higher than those that did the regression
competition, on the classification problem. Using a permutation test,
this corresponds to a significant difference in medians. Similarly the
results show that students who did the regression challenge, performed
better on these exam questions.

Figure \ref{fig:ETC5242} shows the results for students ETC2420/5242.
The boxplots suggest that the students who participated in the challenge
performed relatively better than those that didn't on the regression
question than expected given their total exam performance.

@@

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/ETC5242-1.pdf}
\caption{\label{fig:ETC5242} Performance for regression question
relative to total exam score for students who did and didn't do the
regression data competition in ETC2420/5242.}
\end{figure}

Based on the median, the students who participated in the kaggle
challenge scored 0.09 higher than those that didn't, a median of 1.01 in
comparison to 0.92. Using a permutation test, this corresponds to a
significant difference in medians, with \(p\)-value of 0.015.

@@

Figure \ref{fig:crossperformances} presents students' scores for
classification and regression questions. A score over 1 is considered
outperformance (relative to the expectation). Quarters one and three
include students that underperform or outperform on both types of
questions, respectively. In both cases the number of students that
participated in the classification competition is very close to the
number of students that participated in the regression competition
(excluding a few regression students on the border of score 1). Students
in quarters two and four outperform on one type of questions but not on
the other type. We can see that more regression students outperform on
regression questions than classification students (12 vs. 7). Similarly,
classification students do better on classification questions (11 vs.
3). This is another evidence towards positive influence of the data
competition on student's performances.

@@

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/numsubmition-1.pdf}
\caption{\label{fig:numsubmition} Scatterplots of the exam performance
(a-c) and competition performance (d-f) by number of prediction
submissions, for the three student groups. The relationships with exam
performance are weak. For the MAST90083 and ETC5242 regression
competitions, a clear pattern is that predictions improved substantially
with more submissions. (House price in ETC5242 were divided by 100,000,
explaining the difference in magnitude of error between two
competitions.)}
\end{figure}

@@

Figure \ref{fig:likert} shows the survey responses related to the kaggle
competition, for MAST90083 and ETC5242. The response rate for MAST90083
was 55\%, with 34 of 61 students completing the survey. The response
rate for ETC5242 was 50\%, 17 students out of 34 completed the survey.
Overwhelmingly, students reported that they found the competition
interesting and helpful for their learning in the course.

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/likert-1.pdf}
\caption{\label{fig:likert}Summary of responses to survey of kaggle
competition participants. Overwhelmingly the response to the competition
was positive in both classes, especially the questions on enjoyment and
engagement in the class, and obtaining practical experience. (Table 2
lists the questions.)}
\end{figure}

@@

After collecting the survey from the students we were told that the
questions about students engagement were positively worded. This may
have potentially led to some bias. An improved way could be to ask
directly about student's engagement, e.g. ``How would you rate your
level of engagement in this course?'' with set answer options of ``not
at all engaged'' -- up to ``extremely engaged'' with several choices in
between. Another improvement could be asking ETC2420 students that
didn't take part in the competition about their level of engagement and
compare the answers with other students of ETC5242. We acknowledge that
the differences in the engagement levels may not necessarily be a result
of participation in the competition but it is still an interesting
aspect.

@@

Creating a new competition is surprisingly easy. Information on setting
up a Kaggle InClass challenge is available on the service's web site
(\url{https://www.kaggle.com/about/inclass/overview}). There is a setup
wizard for step-by-step guidance on getting your competition underway.
We have created a short video illustrating the steps to establish a new
competition, available on the web (\url{https://www.youtube.com/XXXX}).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the full set is not available to the students, to avoid cheating.
\item
  the data is not too easy, or too hard, to model so that there is some
  discriminatory power in the results.
\item
  data should be relatively clean, to the point where the instructor has
  tested that a model can be fitted.
\item
  contains some challenges, that make standard off-the-shelf modeling
  less successful, like different variable types that need processing or
  transforming, some outliers, large number of variables.
\item
  if it is a classification challenge, it will work better with
  relatively balanced classes, because the overall accuracy is the
  easiest metric to use.
\end{enumerate}

Choosing the metric upon which to evaluate the model is another
decision. Our advice is to keep it simple, so you, and the students, can
understand the student scores. If you are running a regression
challenge, then ``Root Mean Squared Error (RMSE)''. It is is a
classification challenge, then ``Categorization Accuracy'', the
percentage correct is reasonable.

@@

The competition needs to run without any intervention from the
instructor. The instructor can monitor students progress: the number of
submissions, student scores and even the uploaded data at any time. When
the competition ends the ``Leaderboard'' page provides a list of
students ordered by the final score. It also provides all the scores
from all past submissions (under ``Raw Data'' on ``Public
Leaderboard'').

It may be recommended to limit students to one submission per day. It
encourages students to think about more efficient improvement of their
model before the next submission. It also prevents the student spending
too much time building and submitting models. Some students will become
so engaged in the competition that they might neglect their other
coursework. About halfway through the coompetition, students might be
allowed to form teams, to learn how averaging models can boost
performance.

@@

This paper has described the setup and results of an experiment to
examine the effectiveness of data competitions on student learning,
using Kaggle InClass as the vehicle for conducting the competition. The
experiment was conducted in the classroom settings as part of the normal
teaching of the courses, which imposes limitations on the design.
However, with both MAST90083 and ETC2420/5242, either a randomized
assignment of students to two topic groups, or a control group, was
possible, which enabled comparing performance.

The primary finding, is that participating in a data challenge
competition, produces a statistically significant improvement in the
learning of the topic, although the effect size is small. Secondarily,
anecdotally, the competitions enhanced interest and engagement in the
course.

@@

This document was produced in R \citep{R} with the package knitr
\citep{knitr}. Data cleaning was conducted using tidyr \citep{tidyr},
dplyr \citep{dplyr} and plots were made with ggplot2 \citep{ggplot2}.
The materials to reproduce the work are available at
\url{https://github.com/XXX}.

>>>>

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

%% load any required packages here

\usepackage{setspace}\doublespacing

@@

Kaggle \citep{kaggle} is a platform for predictive modelling and
analytics competitions where participants compete to produce the best
predictive model for a given data set. It is well-known for its
competitions (e.g. \citet{darkmatter}), some of which come with rich
monetary prizes (e.g. \citet{heritage}). There are also learning
competitions \citep{kagglelearn}, designed to help novices hone their
data mining skills. Winners are typically expected to share their code,
and occasionally newly emerged algorithms are introduced to the broad
community, for example, deep neural networks \citep{HintonDahi12} and
XGBoost \citep{ChenXGBoost}.

@@

A competition, like any other active learning method, that used for
assessment, has its advantages and disadvantages. It brings the `game'
feeling, increases the interest level among students and motivates for
higher performance (\citet{Shindler09}, p 105). However, it may have
negative influence if constructed poorly. Among the negative influences
are increased stress and anxiety, induced by fearing a low ranking,
failure or technology barriers. In addition, students may invest a
disproportionate amount of time and effort into competition. Despite
some received criticism, a properly set competition can benefit the
students greatly. The competition should be relatively short in duration
to avoid consuming undue energy. Students should be clear about the
rules and the goal. They should be properly rewarded and most important,
feel that they have reasonable chance to win or achieve high mark
\citep{Shindler09}.

\citet{Shelley08} raised the need for more quantitative and statistical
analysis of evidence in science education. This paper contributes to
this call by offering statistical analysis of the effects on learning of
classroom data competitions.

@@

The experiment was conducted during Semester 2 2017. Data was collected
during two classes, one at the University of Melbourne (Computational
statistics and data mining ,MAST90083, denoted as CSDM), and one at
Monash University (Statistical thinking, ETC2420/5242, denoted as ST).

@@

Computational Statistics and Data Mining (CSDM) is designed for
postgraduate level students with math, statistics, information
technology or actuarial backgrounds. It covers modelling both continuous
(regression) and categorical (classification) response variables. The 63
students were randomized into one of two kaggle competitions, one
focused on regression (R) and the other classification (C). (One of the
63 students elected not to take part in the competition, and another
student didn't sit the exam, producing a final sample size of 61.) This
setup mimics randomized control trials, which is the gold standard, in
experiment design (\citet{Shelley09CH1}, ch.~1). Students built
prediction models and made submissions individually for 16 days, and
then were allowed to form groups to compete for another 7 days.

The reason for this strategy was first to motivate each of the students
to think about modelling and be actively engaged in the competitions
through individual submission. The lecturer allowed participants to
create groups towards the end of the competition to illustrate the
advantages of group work and ensemble models. Another reason for this
approach was the university policy, requiring a strategy to assign
students individually in group assignments. When the team members
develop the model together, it is quite difficult to accurately assess
the individual contribution of each student. The individual submissions
helped to encourage each student to engage in the modelling process. In
addition, it helped to assess the individual component of the final
score for the competition.

Table 1 compares the summary statistics for the two groups. We examine
the percentage correct overall on the final exam for the different
groups and the scores the students received for the second assignment.
The second assignment examined students knowledge about computational
methods, unrelated to the classification and regression methods. The two
groups statistics are similar.

\begin{table}[h]
\begin{center}
\begin{tabular}{l l | r | r | r r r r r}\hline
Score & Comp. group & N & Mean (SD)   & Min  &  Q1 & Median  &  Q3  & Max \\\hline
Exam   & Regression      & 30 &  66.9 (14.1)  &   36.0 &  57.5 &   69.2 &  77.5 &   88.5\\
       & Classification & 31 &  62.1 (17.2)  &   20.0  & 52.2  &  68.0   &  75.2 &  85.5\\\hline

2$^{nd}$ assignment & Regression      &    30  & 8.2 (2.2)  &  0.0  & 7.0 &  8.4   &  9.9  &   10.0 \\
                    & Classification  &    31  & 8.0 (3.0)  &  0.0  & 7.8  &  9.3   &  9.9  &   10.0 \\\hline

\end{tabular}
\caption{Computational statistics and data mining: Summary statistics of the exam score (out of 100) and the second assignment (out of 10) for the two competition groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

Statistical Thinking (ST), covers regression, but not classification,
and has a mix of undergraduate and postgraduate students. Only the 34
postgraduate (ST-PG) students were required to participate in the kaggle
competition, and competed in the regression (R) challenge. This was run
independently from the CSDM competition. The 145 undergraduate (ST-UG)
students were used as comparison for examining performance of the
postgraduate students. Although, it may be surprising at first glance,
the undergraduate students provide a reasonable comparison for the
graduate students. The class is taught to both cohorts simultaneously.
The entry requirements to the Bachelors of Commerce at Monash is high,
and these students have strong mathematics backgrounds. In the years
prior to this experiment, the undergraduate scores on the final exam are
indistinguishable from those of the graduate students. The competition
ran for one month. Students formed their own teams of 2-4 members to
compete.

Table 2 shows the summary statistics of the exam scores for the 34
postgraduate (ST-PG) students and for the 145 undergraduate (ST-UG)
students. The mean and the median scores of postgraduate students are a
bit lower than the corresponding scores of undergraduate students.
However, the interquartile range is similar. Interestingly, the highest
exam score was received by an undergraduate student.

\begin{table}[h]
\begin{center}
\begin{tabular}{l l | r | r | r r r r r}\hline
Score & Competition group & N & Mean (SD)   & Min  &  Q1 & Median  &  Q3  & Max \\\hline
Exam   & ST-PG      & 34  & 61.8 (12.2)  &  35.0 & 57.2  &   61.0  & 72.2  & 81.5 \\
       & ST-UG         & 145 & 63.0 (15.4)  &   0.0 & 56.0  &   64.0  & 73.0   & 90.0\\\hline
\end{tabular}
\caption{Statistical thinking: Summary statistics of the exam score (out of 100) for the two groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

Figure \ref{fig:Monash_quiz} displays the quiz scores taken by the
undergraduate (ST-UG) and postgraduate (ST-PG) students, as a beeswarm
plot \citep{beeswarm}. The distribution of quiz scores are similar,
except that undergraduates tend to have a longer lower tail.

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/Monash_quiz-1.pdf}
\caption{\label{fig:Monash_quiz}Beeswarm plots of undergraduate (ST-UG)
and postgraduate (ST-PG) students quizzes during the semester. The
scores are very similar, although undergraduates have a longer lower
tail.}
\end{figure}

@@

Figure \ref{fig:MAST90083} shows the data collected in CSDM. Performance
is plotted against type of question, separately for the competition they
completed. The difference in median scores indicates performance
improvement. Students who completed the classification competition
(left) performed relatively better on the classification questions than
the regression questions in the final exam. Conversely, students who
participated in the regression competition performed relatively better
on the regression questions.

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/Melb-1.pdf}
\caption{\label{fig:MAST90083} Boxplots of performance on regression and
classification questions in the final exam, by type of data competition
completed in CSDM. Students generally performed better on the questions
corresponding to the competition they participated in.}
\end{figure}

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|r|r|}\hline
Question Set & Median difference & Permutation $p$-value \\\hline
Classification & 0.250 & 0.033\\
Regression & 0.104 & 0.000\\\hline
\end{tabular}
\caption{Comparison of median difference in performance by competition group, for CSDM students, using permutation tests. Both sets of medians are significantly different, indicating improved scores for questions on the topic related to the Kaggle competition.}
\end{center}
\label{tab:Melb_Perm}
\end{table}

Table 3 shows the results of permutation testing of median difference
between the groups. Generally the results support the competition
improved performance. Students who participated in the kaggle challenge
for classification scored higher than those that did the regression
competition, on the classification problem. Using a permutation test,
this corresponds to a significant difference in medians. Similarly the
results show that students who did the regression challenge, performed
better on these exam questions.

Figure \ref{fig:ETC5242} shows the results for ST students. The boxplots
suggest that the students who participated in the challenge performed
relatively better than those that didn't on the regression question than
expected given their total exam performance.

@@

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/ETC5242-1.pdf}
\caption{\label{fig:ETC5242} Performance for regression question
relative to total exam score for students who did and didn't do the
regression data competition in Statistical thinking.}
\end{figure}

Based on the median, the students who participated in the kaggle
challenge scored 0.09 higher than those that didn't, a median of 1.01 in
comparison to 0.92. Using a permutation test, this corresponds to a
significant difference in medians, with \(p\)-value of 0.02.

@@

Figure \ref{fig:crossperformances} presents students' scores for
classification and regression questions. A score over 1 is considered as
outperform (relative to the expectation). Quarters one and three include
students that underperform or outperform on both types of questions,
respectively. In both cases the number of students that participated in
the classification competition is very close to the number of students
that participated in the regression competition (excluding a few
regression students on the border of score 1). Students in quarters two
and four outperform on one type of questions but not on the other type.
We can see that more regression students outperform on regression
questions than classification students (12 vs. 7). Similarly,
classification students do better on classification questions (11 vs.
3). This is another evidence towards positive influence of the data
competition on student's performances.

@@

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/numsubmition-1.pdf}
\caption{\label{fig:numsubmition} Scatterplots of the exam performance
(a-c) and competition performance (d-f) by number of prediction
submissions, for the three student groups. The relationships with exam
performance are weak. For the CSDM and ST-PG regression competitions, a
clear pattern is that predictions improved substantially with more
submissions. (House price in ST-PG were divided by 100,000, explaining
the difference in magnitude of error between two competitions.)}
\end{figure}

@@

Figure \ref{fig:likert} shows the survey responses related to the kaggle
competition, for CSDM and ST-PG. The response rate for CSDM was 55\%,
with 34 of 61 students completing the survey. The response rate for
ST-PG was 50\%, 17 students out of 34 completed the survey.
Overwhelmingly, students reported that they found the competition
interesting and helpful for their learning in the course.

\begin{figure}
\centering
\includegraphics{paper-kaggle_files/figure-latex/likert-1.pdf}
\caption{\label{fig:likert}Summary of responses to survey of kaggle
competition participants. Overwhelmingly the response to the competition
was positive in both classes, especially the questions on enjoyment and
engagement in the class, and obtaining practical experience. (Table 4
lists the questions.)}
\end{figure}

@@

After collecting the survey from the students we were told that the
questions about students engagement were positively worded. This may
have potentially led to some bias. An improved way could be to ask
directly about student's engagement, e.g. ``How would you rate your
level of engagement in this course?'' with set answer options of ``not
at all engaged'' -- up to ``extremely engaged'' with several choices in
between. Another improvement could be asking ST-UG students that didn't
take part in the competition about their level of engagement and compare
the answers with other students of ST-PG. We acknowledge that the
differences in the engagement levels may not necessarily be a result of
participation in the competition but it is still an interesting aspect.

@@

Creating a new competition is surprisingly easy. Information on setting
up a Kaggle InClass challenge is available on the service's web site
(\url{https://www.kaggle.com/about/inclass/overview}). There is a setup
wizard for step-by-step guidance on getting your competition underway.
We have created a short video illustrating the steps to establish a new
competition, available on the web
(\url{https://www.youtube.com/watch?v=tqbps4vq2Mc&t=32s}).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the full set is not available to the students, to avoid plagiarism and
  use of unauthorized assistance.
\item
  the data is not too easy, or too hard, to model so that there is some
  discriminatory power in the results.
\item
  data should be relatively clean, to the point where the instructor has
  tested that a model can be fitted.
\item
  contains some challenges, that make standard off-the-shelf modeling
  less successful, like different variable types that need processing or
  transforming, some outliers, large number of variables.
\item
  if it is a classification challenge, it will work better with
  relatively balanced classes, because the overall accuracy is the
  easiest metric to use.
\end{enumerate}

The data sets used in our competitions can be shared with other
instructors by request.

Choosing the metric upon which to evaluate the model is another
decision. Our advice is to keep it simple, so you, and the students, can
understand the student scores. If you are running a regression
challenge, then the ``Root Mean Squared Error (RMSE)'' is a good choice.
If it is a classification challenge, then ``Categorization Accuracy'',
the percent of correct classifications, is reasonable.

@@

The competition needs to run without any intervention from the
instructor. Exception is, of course, an academic discussion motivated by
the competition between the teaching team and the students. For example,
a discussion about different models, their advantages and limitations.
The instructor can monitor students progress: the number of submissions,
student scores and even the uploaded data at any time. When the
competition ends the ``Leaderboard'' page provides a list of students
ordered by the final score. It also provides all the scores from all
past submissions (under ``Raw Data'' on ``Public Leaderboard'').

It may be recommended to limit students to one submission per day. It
encourages students to think about more efficient improvement of their
model before the next submission. It also prevents the student spending
too much time building and submitting models. Some students will become
so engaged in the competition that they might neglect their other
coursework. About halfway through the competition, students might be
allowed to form teams, to learn how averaging models can boost
performance.

@@

This paper has described an experiment to examine the effectiveness of
data competitions on student learning, using Kaggle InClass as the
vehicle for conducting the competition. The primary finding, is that
participating in a data challenge competition, produces a statistically
significant improvement in the learning of the topic, although the
effect size is small. Secondarily, the competitions enhanced interest
and engagement in the course.

Because the experiment was conducted in the classroom setting as part of
the normal teaching of the courses, which imposed limitations on the
design. However, performance comparison was enabled in CSDM by a
randomized assignment of students to two topic groups, and in ST by
using a control group. In CDSM, the group sizes were relatively small,
approximately 30 students per group. For ST the control group was the
undergraduate students that took class. Using undergraduate students as
a control group for graduate students may be surprising. However, the
experience of teaching this subject over several years and some
statistical comparison of the two groups justifies the approach.
Undergraduate students performance in other tasks and exam questions,
not relevant to the competition, was equivalent to the postgraduate
students cohort.

@@

This document was produced in R \citep{R} with the package knitr
\citep{knitr}. Data cleaning was conducted using tidyr \citep{tidyr},
dplyr \citep{dplyr} and plots were made with ggplot2 \citep{ggplot2}.
The materials to reproduce the work are available at
\url{https://github.com/dicook/paper-quoll}.

### 2021-01-04 Submission3/paper-kaggle.Rmd
Submission 3 created

<!-- Unblinded verssion -->
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of Melbourne (Computational statistics and data mining ,MAST90083, denoted as CSDM), and one at Monash University (Statistical thinking, ETC2420/5242, denoted as ST).

>>>>

<!-- Unblinded verssion -->
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of Melbourne (Computational statistics and data mining, MAST90083, denoted as CSDM), and one at Monash University (Statistical thinking, ETC2420/5242, denoted as ST).

### 2021-01-04 Submission3/paper-kaggle.tex
Submission 3 created

The experiment was conducted during Semester 2 2017. Data was collected
during two classes, one at the University of Melbourne (Computational
statistics and data mining ,MAST90083, denoted as CSDM), and one at
Monash University (Statistical thinking, ETC2420/5242, denoted as ST).

@@

Based on the median, the students who participated in the kaggle
challenge scored 0.09 higher than those that didn't, a median of 1.01 in
comparison to 0.92. Using a permutation test, this corresponds to a
significant difference in medians, with \(p\)-value of 0.02.

>>>>

The experiment was conducted during Semester 2 2017. Data was collected
during two classes, one at the University of Melbourne (Computational
statistics and data mining, MAST90083, denoted as CSDM), and one at
Monash University (Statistical thinking, ETC2420/5242, denoted as ST).

@@

Based on the median, the students who participated in the kaggle
challenge scored 0.09 higher than those that didn't, a median of 1.01 in
comparison to 0.92. Using a permutation test, this corresponds to a
significant difference in medians, with \(p\)-value of 0.024.
